---
title: "p8105_HW1_rat2134"
author: "Robert Tumasian"
date: "9/12/2019"
output: github_document
---

```{r}
###Loading required packages
library(tidyverse)
```

```{r}
###PROBLEM 1
set.seed(1234) #Set seed to guarantee reproducible results

#Create data set
Problem_1_df=tibble(
  x = rnorm(n = 8, mean = 0, sd = 1),
  vec_logical = c(x > 0),
  vec_char = c("a","b","c","d","e","f","g","h"),
  vec_factor = factor(c("Red","Red","Blue","Red","Green","Green","Red","Red"))
)

#Obtaining the mean of each variable in data set
mean(pull(Problem_1_df, x)) 
mean(pull(Problem_1_df, vec_logical)) #half of the elements of x are greater than 0 and half are below 0

mean(pull(Problem_1_df, vec_char)) 

mean(pull(Problem_1_df, vec_factor)) 

```

```{r, eval=FALSE}
#Converting logical, character, and factor vectors to numeric vectors
as.numeric(pull(Problem_1_df, vec_logical)) #TRUE values are converted to 1s and FALSE values are converted to 0s

as.numeric(pull(Problem_1_df, vec_char)) #Vector becomes all NAs, because character vectors cannot be enumerated

as.numeric(pull(Problem_1_df, vec_factor)) #Levels are enumerated alphabetically ("Blue" = 1, "Green" = 2, and "Red" = 3)

```

It makes sense that we are only able to take the mean of vectors that are able to be enuerated (logical and factor vectors), but not vectors that are unable to be enumerated (character vectors).

```{r}
#Converting logical vector to numeric, and multiplying the random sample by the result
as.numeric(pull(Problem_1_df, vec_logical)) * pull(Problem_1_df, x)

#Converting the logical vector to a factor, and multiplying the random sample by the result
as.factor(pull(Problem_1_df, vec_logical)) * pull(Problem_1_df, x)

#Converting the logical vector to a factor and then converting the result to numeric, and multiplying the random sample by the result
as.numeric(as.factor(pull(Problem_1_df, vec_logical))) * pull(Problem_1_df, x)

```

```{r}
###PROBLEM 2
set.seed(12345) #Set seed to guarantee reproducible results

#Create data set
Problem_2_df=tibble(
  x = rnorm(n = 500, mean = 0, sd = 1),
  y = rnorm(n = 500, mean = 0, sd = 1),
  vec_logical = c(x + y > 1),
  vec_numeric = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
)
```

This data set has `r nrow(Problem_2_df)` rows and `r ncol(Problem_2_df)` columns. The mean, median, and standard deviation of x are `r mean(pull(Problem_2_df, x))`, `r median(pull(Problem_2_df, x))`, and `r sd(pull(Problem_2_df, x))`, respectively. Lastly, the proportion of cases for which `x + y > 1` is `r (length(which(pull(Problem_2_df, vec_logical)) == "TRUE") / length(pull(Problem_2_df, vec_logical)))`. Therefore, 26% of the `x + y` values are greater than 1.